---
title: "Case Study EDA"
author: "Isiah Perine"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(e1071)
library(class)
library(caret)
```


```{r}
# Reading in the data

data = read.csv("CaseStudy1-data.csv")
data
```

```{r}
# checking for NA's

sum(is.na(data))
```

```{r}
# manipulating variables 

data$Attrition = as.factor(data$Attrition)
data$BusinessTravel = as.factor(data$BusinessTravel)
data$Department = as.factor(data$Department)
data$Education = as.factor(data$Education)
data$EducationField = as.factor(data$EducationField)
data$EnvironmentSatisfaction = as.factor(data$EnvironmentSatisfaction)
data$Gender = as.factor(data$Gender)
data$JobInvolvement = as.factor(data$JobInvolvement)
data$JobLevel = as.factor(data$JobLevel)
data$JobRole = as.factor(data$JobRole)
data$JobSatisfaction = as.factor(data$JobSatisfaction)
data$MaritalStatus = as.factor(data$MaritalStatus)
data$Over18 = as.factor(data$Over18)
data$OverTime = as.factor(data$OverTime)
data$PerformanceRating = as.factor(data$PerformanceRating)
data$RelationshipSatisfaction = as.factor(data$RelationshipSatisfaction)
data$StockOptionLevel = as.factor(data$StockOptionLevel)
data$WorkLifeBalance = as.factor(data$WorkLifeBalance)

data
```


# EDA PORTION


```{r}
# Attrition vs Age

summary(data$Age)

data$AgeF = cut(data$Age, 
                      breaks = c(17, 35, 61), 
                      labels = c("Young", "Old"))

attrition_percentage = data %>%
  group_by(AgeF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = AgeF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Age") +
  ylab("Percentage") +
  xlab("Age")
```

*Young seems to have a higher percentage*


```{r}
# Attrition vs Business Travel

attrition_percentage = data %>%
  group_by(BusinessTravel, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = BusinessTravel, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Business Travel") +
  ylab("Percentage") +
  xlab("Business Travel")
```

*Travel Frequently seems to have the highest percentage*


```{r}
# Attrition vs Daily Rate

summary(data$DailyRate)

data$DailyRateF = cut(data$DailyRate, 
                      breaks = c(100, 817, 1500), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(DailyRateF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = DailyRateF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Daily Rate") +
  ylab("Percentage") +
  xlab("Daily Rate")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Department

attrition_percentage = data %>%
  group_by(Department, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = Department, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Department") +
  ylab("Percentage") +
  xlab("Department")
```

*Sales seems to have the highest percentage*


```{r}
# Attrition vs Distance From Home

summary(data$DistanceFromHome)

data$DistanceFromHomeF = cut(data$DistanceFromHome, 
                      breaks = c(0, 7, 30), 
                      labels = c("Close", "Far"))

attrition_percentage = data %>%
  group_by(DistanceFromHomeF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = DistanceFromHomeF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Distance From Home") +
  ylab("Percentage") +
  xlab("Distance From Home")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Education

attrition_percentage = data %>%
  group_by(Education, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = Education, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Education") +
  ylab("Percentage") +
  xlab("Education")
```

*Below college seems to have the highest percentage. There also seems to be a decreasing pattern in percentage as education level increases*


```{r}
# Attrition vs Education Field

attrition_percentage = data %>%
  group_by(EducationField, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = EducationField, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Education Field") +
  ylab("Percentage") +
  xlab("Education Field")
```

*Human Resources seems to have the highest percentage* 


```{r}
# Attrition vs Environment Satisfaction

attrition_percentage = data %>%
  group_by(EnvironmentSatisfaction, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = EnvironmentSatisfaction, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Environment Satisfaction") +
  ylab("Percentage") +
  xlab("Environment Satisfaction")
```

*Low environment satisfaction seems to have the highest percentage*


```{r}
# Attrition vs Gender

attrition_percentage = data %>%
  group_by(Gender, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
attrition_percentage

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = Gender, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Gender") +
  ylab("Percentage") +
  xlab("Gender")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Hourly Rate

summary(data$HourlyRate)

data$HourlyRateF = cut(data$HourlyRate, 
                      breaks = c(29, 66, 101), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(HourlyRateF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = HourlyRateF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Hourly Rate") +
  ylab("Percentage") +
  xlab("Hourly Rate")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Job Involvement

attrition_percentage = data %>%
  group_by(JobInvolvement, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
attrition_percentage

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = JobInvolvement, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Job Involvement") +
  ylab("Percentage") +
  xlab("Job Involvement")
```

*Low job involvement seems to have the highest percentage. There also seems to be a decreasing pattern in percentage as job involvement increases*


```{r}
# Attrition vs Job Level

attrition_percentage = data %>%
  group_by(JobLevel, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = JobLevel, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Job Level") +
  ylab("Percentage") +
  xlab("Job Level")
```

*The lowest job level seems to have the highest percentage. There does not seem to be a pattern in percentage as job level increases*


```{r}
# Attrition vs Job Role

attrition_percentage = data %>%
  group_by(JobRole, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
attrition_percentage

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = JobRole, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  ggtitle("Percentage of Attrition by Job Role") +
  ylab("Percentage") +
  xlab("Job Role")
```

*Sales Representative seems to have the highest percentage*


```{r}
# Attrition vs Job Satisfaction

attrition_percentage = data %>%
  group_by(JobSatisfaction, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = JobSatisfaction, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Job Satisfaction") +
  ylab("Percentage") +
  xlab("Job Satisfaction")
```

*Low job satisfaction seems to have the highest percentage. There also seems to be a decreasing pattern in percentage as job satisfaction increases*


```{r}
# Attrition vs Marital Status

attrition_percentage = data %>%
  group_by(MaritalStatus, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = MaritalStatus, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Marital Status") +
  ylab("Percentage") +
  xlab("Marital Status")
```

*Single marital status seems to have the highest percentage. There also seems to be a decreasing pattern in percentage as marital status goes from single -> married -> divorced*


```{r}
# Attrition vs Monthly Income

summary(data$MonthlyIncome)

data$MonthlyIncomeF = cut(data$MonthlyIncome, 
                      breaks = c(1080, 4946, 20000), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(MonthlyIncomeF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = MonthlyIncomeF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Monthly Income") +
  ylab("Percentage") +
  xlab("Monthly Income")
```

*Low monthly income seems to have a higher percentage*


```{r}
# Attrition vs Monthly Rate

summary(data$MonthlyRate)

data$MonthlyRateF = cut(data$MonthlyRate, 
                      breaks = c(2093, 14074, 26998), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(MonthlyRateF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = MonthlyRateF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Monthly Rate") +
  ylab("Percentage") +
  xlab("Monthly Rate")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Number of Companies Worked At

summary(data$NumCompaniesWorked)

data$NumCompaniesWorkedF = cut(data$NumCompaniesWorked, 
                      breaks = c(-1, 2, 10), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(NumCompaniesWorkedF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = NumCompaniesWorkedF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Number of Companies Worked") +
  ylab("Percentage") +
  xlab("Number of Companies Worked At")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Over 18

attrition_percentage = data %>%
  group_by(Over18, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage, 
       aes(x = Attrition, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Over 18") +
  ylab("Percentage") +
  xlab("Attrition")
```

*All individuals in the data set are over 18. All this plot tells me is that about 16% of the individuals in the data set fall into the 'yes' category for attrition*


```{r}
# Attrition vs Over Time

attrition_percentage = data %>%
  group_by(OverTime, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
attrition_percentage

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = OverTime, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Over Time") +
  ylab("Percentage") +
  xlab("Over Time")
```

*Over time workers seems to have a higher percentage*


```{r}
# Attrition vs Percent Salary Hike

summary(data$PercentSalaryHike)

data$PercentSalaryHikeF = cut(data$PercentSalaryHike, 
                      breaks = c(10, 14, 26), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(PercentSalaryHikeF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = PercentSalaryHikeF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Percent Salary Hike") +
  ylab("Percentage") +
  xlab("Percent Salary Hike")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Performance Rating

attrition_percentage = data %>%
  group_by(PerformanceRating, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = PerformanceRating, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Performance Rating") +
  ylab("Percentage") +
  xlab("Performance Rating")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Relationship Satisfaction

attrition_percentage = data %>%
  group_by(RelationshipSatisfaction, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = RelationshipSatisfaction, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Relationship Satisfaction") +
  ylab("Percentage") +
  xlab("Relationship Satisfaction")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Standard Hours

attrition_percentage = data %>%
  group_by(StandardHours, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage, 
       aes(x = Attrition, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Standard Hours") +
  ylab("Percentage") +
  xlab("Attrition")
```

*All individuals in the data set have 80 standard hours. All this plot tells me is that about 16% of the individuals in the data set fall into the 'yes' category for attrition*


```{r}
# Attrition vs Stock Option Level

attrition_percentage = data %>%
  group_by(StockOptionLevel, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = StockOptionLevel, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Stock Option Level") +
  ylab("Percentage") +
  xlab("Stock Option Level")
```

*0 and 3 stock option levels seem to have the highest percentage*


```{r}
# Attrition vs Total Working Years

summary(data$TotalWorkingYears)

data$TotalWorkingYearsF = cut(data$TotalWorkingYears, 
                      breaks = c(-1, 10, 40), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(TotalWorkingYearsF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = TotalWorkingYearsF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Total Working Years") +
  ylab("Percentage") +
  xlab("Total Working Years")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Training Times Last Years

summary(data$TrainingTimesLastYear)

data$TrainingTimesLastYearF = cut(data$TrainingTimesLastYear, 
                      breaks = c(-1, 3, 6), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(TrainingTimesLastYearF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = TrainingTimesLastYearF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Training Times Last Year") +
  ylab("Percentage") +
  xlab("Training Times Last Year")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Work Life Balance

attrition_percentage = data %>%
  group_by(WorkLifeBalance, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = WorkLifeBalance, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Work Life Balance") +
  ylab("Percentage") +
  xlab("Work Life Balance")
```

*Bad work life balance seems to have the highest percentage. There also seems to be a decreasing pattern in percentage as work life balance gets better*


```{r}
# Attrition vs Years At Company

summary(data$YearsAtCompany)

data$YearsAtCompanyF = cut(data$YearsAtCompany, 
                      breaks = c(-1, 5, 40), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(YearsAtCompanyF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = YearsAtCompanyF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Years At Company") +
  ylab("Percentage") +
  xlab("Years At Company")
```

*Low years at company seems to have a higher percentage*


```{r}
# Attrition vs Years In Current Role

summary(data$YearsInCurrentRole)

data$YearsInCurrentRoleF = cut(data$YearsInCurrentRole, 
                      breaks = c(-1, 3, 18), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(YearsInCurrentRoleF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = YearsInCurrentRoleF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Years In Current Role") +
  ylab("Percentage") +
  xlab("Years In Current Role")
```

*Low years in current role seems to have a higher percentage*


```{r}
# Attrition vs Years Since Last Promotion

summary(data$YearsSinceLastPromotion)

data$YearsSinceLastPromotionF = cut(data$YearsSinceLastPromotion, 
                      breaks = c(-1, 1, 15), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(YearsSinceLastPromotionF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = YearsSinceLastPromotionF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Years Since Last Promotion") +
  ylab("Percentage") +
  xlab("Years Since Last Promotion")
```

*Not much of a percentage difference per group*


```{r}
# Attrition vs Years With Current Manager

summary(data$YearsWithCurrManager)

data$YearsWithCurrManagerF = cut(data$YearsWithCurrManager, 
                      breaks = c(-1, 3, 17), 
                      labels = c("Low", "High"))

attrition_percentage = data %>%
  group_by(YearsWithCurrManagerF, Attrition) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(attrition_percentage[attrition_percentage$Attrition == "Yes", ], 
       aes(x = YearsWithCurrManagerF, y = percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  ggtitle("Percentage of Attrition by Years With Current Manager") +
  ylab("Percentage") +
  xlab("Years With Current Manager")
```

*Not much of a percentage difference per group*



# All of the variables with visual evidence of an impact to attrition

# Age
# BusinessTravel
# Department
# Education
# EducationField
# EnvironmentSatisfaction
# JobInvolvement
# JobLevel
# JobRole
# JobSatisfaction
# MaritalStatus
# MonthlyIncome
# OverTime
# StockOptionLevel
# WorkLifeBalance
# YearsAtCompany
# YearsInCurrentRole

```{r}
# Chi-squared test to check for significantly correlated variables 

chisq.test(table(data$Attrition, data$Age)) #p-value = 0.0004559
chisq.test(table(data$Attrition, data$BusinessTravel)) #p-value = 0.04993
chisq.test(table(data$Attrition, data$Department)) #p-value = 0.009424
chisq.test(table(data$Attrition, data$Education)) #p-value = 0.6243
chisq.test(table(data$Attrition, data$EducationField)) #p-value = 0.2682
chisq.test(table(data$Attrition, data$JobInvolvement)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$JobLevel)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$JobRole)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$JobSatisfaction)) #p-value = 0.01115
chisq.test(table(data$Attrition, data$MaritalStatus)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$MonthlyIncome)) #p-value = 0.5485
chisq.test(table(data$Attrition, data$OverTime)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$StockOptionLevel)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$WorkLifeBalance)) #p-value = 0.002495
chisq.test(table(data$Attrition, data$YearsAtCompany)) #p-value < 0.0001
chisq.test(table(data$Attrition, data$YearsInCurrentRole)) #p-value = 0.0007929
```

# The variales with the most significant correlation

# 1.) JobInvolvement
# 2.) JobLevel
# 3.) JobRole
# 4.) MaritalStatus
# 5.) OverTime
# 6.) StockOptionLevel
# 7.) YearsAtCompany

# I can make 35 models with 3 of the 7 varibles I have listed

```{r}
# calculating mean Accuracy, Sensitivity, and Specificity

AccHolder = numeric(100)
SensHolder = numeric(100)
SpecHolder = numeric(100)

for (seed in 1:100)
{
set.seed(seed)
trainIndices = sample(seq(1:length(data$Attrition)), round(.7*length(data$Attrition)))
train = data[trainIndices, ]
test = data[-trainIndices, ]
model = naiveBayes(train[, c("OverTime", "StockOptionLevel", "JobRole")], train$Attrition)
CM = confusionMatrix(table(predict(model, test[, c("OverTime", "StockOptionLevel", "JobRole")]), test$Attrition))
AccHolder[seed] = CM$overall[1]
SensHolder[seed] = CM$byClass[1]
SpecHolder[seed] = CM$byClass[2]
}

mean(AccHolder)
mean(SensHolder)
mean(SpecHolder)
```



```{r}
# creating confusion matrices for presentation 

# best
set.seed(1)
trainIndices = sample(seq(1:length(data$Attrition)), round(.7*length(data$Attrition)))
train = data[trainIndices, ]
test = data[-trainIndices, ]
model = naiveBayes(train[, c("MaritalStatus", "OverTime", "StockOptionLevel")], train$Attrition)
confusionMatrix(table(predict(model, test[, c("MaritalStatus", "OverTime", "StockOptionLevel")]), test$Attrition))

# second best
set.seed(1)
trainIndices = sample(seq(1:length(data$Attrition)), round(.7*length(data$Attrition)))
train = data[trainIndices, ]
test = data[-trainIndices, ]
model = naiveBayes(train[, c("JobLevel", "OverTime", "StockOptionLevel")], train$Attrition)
confusionMatrix(table(predict(model, test[, c("JobLevel", "OverTime", "StockOptionLevel")]), test$Attrition))
```


```{r}
# finding the "best" model

AccHolder = numeric(100)
SensHolder = numeric(100)
SpecHolder = numeric(100)

for (seed in 1:100)
{
set.seed(seed)
trainIndices = sample(seq(1:length(data$Attrition)), round(.7*length(data$Attrition)))
train = data[trainIndices, ]
test = data[-trainIndices, ]
model = naiveBayes(train[, c("JobInvolvement", "JobLevel", "OverTime", "YearsAtCompany", "JobRole", "Age", "BusinessTravel", "Education", "MaritalStatus", "StockOptionLevel", "WorkLifeBalance", "YearsInCurrentRole", "EnvironmentSatisfaction", "MonthlyIncome", "DistanceFromHome", "YearsWithCurrManager", "TotalWorkingYears")], train$Attrition)
CM = confusionMatrix(table(predict(model, test[, c("JobInvolvement", "JobLevel", "OverTime", "YearsAtCompany", "JobRole", "Age", "BusinessTravel", "Education", "MaritalStatus", "StockOptionLevel", "WorkLifeBalance", "YearsInCurrentRole", "EnvironmentSatisfaction", "MonthlyIncome", "DistanceFromHome", "YearsWithCurrManager", "TotalWorkingYears")]), test$Attrition))
AccHolder[seed] = CM$overall[1]
SensHolder[seed] = CM$byClass[1]
SpecHolder[seed] = CM$byClass[2]
}

mean(AccHolder)
mean(SensHolder)
mean(SpecHolder)
```



# FINAL MODEL

```{r}
# reading in data

competition = read.csv("CaseStudy1CompSet No Attrition.csv")
competition
```

```{r}
# manipulating variables 

competition$BusinessTravel = as.factor(competition$BusinessTravel)
competition$Department = as.factor(competition$Department)
competition$Education = as.factor(competition$Education)
competition$EducationField = as.factor(competition$EducationField)
competition$EnvironmentSatisfaction = as.factor(competition$EnvironmentSatisfaction)
competition$Gender = as.factor(competition$Gender)
competition$JobInvolvement = as.factor(competition$JobInvolvement)
competition$JobLevel = as.factor(competition$JobLevel)
competition$JobRole = as.factor(competition$JobRole)
competition$JobSatisfaction = as.factor(competition$JobSatisfaction)
competition$MaritalStatus = as.factor(competition$MaritalStatus)
competition$Over18 = as.factor(competition$Over18)
competition$OverTime = as.factor(competition$OverTime)
competition$PerformanceRating = as.factor(competition$PerformanceRating)
competition$RelationshipSatisfaction = as.factor(competition$RelationshipSatisfaction)
competition$StockOptionLevel = as.factor(competition$StockOptionLevel)
competition$WorkLifeBalance = as.factor(competition$WorkLifeBalance)
```


```{r}
# reating predictions

final.model = naiveBayes(data[, c("JobInvolvement", "JobLevel", "OverTime", "YearsAtCompany", "JobRole", "Age", "BusinessTravel", "Education", "MaritalStatus", "StockOptionLevel", "WorkLifeBalance", "YearsInCurrentRole", "EnvironmentSatisfaction", "MonthlyIncome", "DistanceFromHome", "YearsWithCurrManager", "TotalWorkingYears")], data$Attrition)

preds = predict(model, competition[, c("JobRole", "OverTime", "YearsAtCompany")])
preds
```

```{r}
# creating data frame for IDs and predictions

submit = data.frame(ID = competition$ID, Attrition = preds)
submit

yes = subset(submit, submit$Attrition == "Yes")
yes

no = subset(submit, submit$Attrition == "No")
no
```

```{r}
# creating csv with my solutions

write_csv(submit, "Case1PredictionsPerine Attrition.csv")
```


